%===========================================================================
% APPENDIX: ENIGH Data Construction
% For: "The Effect of PROGRESA on Adult Mortality"
% Journal: Journal of Development Economics
%===========================================================================

\clearpage
\appendix

\section{ENIGH Data Construction}
\label{app:enigh}

This appendix documents the complete construction of all outcome variables
drawn from the Encuesta Nacional de Ingresos y Gastos de los Hogares (ENIGH)
used in the mechanisms analysis (Section~\ref{sec:mechanisms}). We use nine
waves: 1992, 1994, 1996, 1998, 2000 (processed in \texttt{01\_enigh\_data.do}
as the early-wave loop) and 2002, 2004, 2005, and 2006 (processed in the
late-wave loop). All spending variables are drawn from \texttt{gastos.dta}
(expenditure records), income variables from \texttt{ingresos.dta}, financial
outflows from \texttt{eroga.dta}, and individual-level characteristics from
\texttt{pob.dta} and \texttt{concen.dta}, all merged on household identifier
\texttt{FOLIO}. A-prefix codes denote expenditure items; J-prefix codes denote
health expenditures; P-prefix codes denote income sources; Q-prefix codes
denote financial outflows.

All quarterly reported amounts (\texttt{GAS\_TRI}, \texttt{ING\_TRI},
\texttt{ERO\_TRI}) are converted to monthly by dividing by three. The 1992
wave reports monetary values in old Mexican pesos (prior to the 1993
redenomination at 1,000:1); these are rescaled by dividing by 1,000 before
all other operations. Section~\ref{app:enigh_deflation} describes the CPI and
exchange-rate deflation applied uniformly after all waves are appended.

%--------------------------------------------------------------------------
\subsection{Food Expenditure Variables}
\label{app:enigh_food}
%--------------------------------------------------------------------------

Food spending is extracted from \texttt{gastos.dta} at the item level, with
each record identified by a catalogue code (\texttt{CLAVE}) composed of a
letter prefix (``A'' for food) and a numeric suffix. An indicator variable is
constructed for each food category based on the numeric range of the A-code;
the monthly household spending in that category equals the indicator times
\texttt{GAS\_TRI/3}. The dataset is then collapsed to the household level by
summing all items within each category.

The final analysis variables are formed by aggregating sub-categories:

\begin{align*}
\texttt{cereals}          &= \texttt{cereals\_d} \times \texttt{GAS\_TRI}/3 \\
\texttt{meat\_dairy}      &= (\texttt{meat\_fish\_seafood\_d} + \texttt{dairy\_d})
                              \times \texttt{GAS\_TRI}/3 \\
\texttt{vegg\_fruit}      &= \texttt{vegg\_fruit\_d} \times \texttt{GAS\_TRI}/3 \\
\texttt{sugar\_fat\_drink}&= (\texttt{oils\_fats\_d} + \texttt{sugar\_d}
                              + \texttt{soft\_drink\_d} + \texttt{desserts\_d})
                              \times \texttt{GAS\_TRI}/3 \\
\texttt{coffe\_spices\_other} &= (\texttt{coffe\_d} + \texttt{specias\_d}
                              + \texttt{others\_d}) \times \texttt{GAS\_TRI}/3 \\
\texttt{outside\_food}    &= (\texttt{takeout\_d} + \texttt{outside\_d})
                              \times \texttt{GAS\_TRI}/3 \\
\texttt{tobacco}          &= \texttt{tobacco\_d} \times \texttt{GAS\_TRI}/3 \\
\texttt{alcohol}          &= \texttt{alcohol\_d} \times \texttt{GAS\_TRI}/3
\end{align*}

\noindent
\texttt{packaged\_food} (pre-packaged despensas), \texttt{baby\_food}, and
\texttt{pet\_food} are constructed analogously. \texttt{packaged\_food\_d} is
set to zero for all 1992--2000 waves; the category only appears in the
catalogue from 2002 onward, identified by the specific codes A209 and A243.

Table~\ref{tab:food_acodes} reports the A-code boundaries for each food
category by wave. The numbering of catalogue items shifts across waves as INEGI
expands or reorganises the catalogue; boundaries were verified against the
official ENIGH catalogue PDFs for each year.

\begin{table}[h!]
\centering
\caption{Food Expenditure A-Code Boundaries by ENIGH Wave}
\label{tab:food_acodes}
\scriptsize
\begin{tabular}{lrrrrrrrrr}
\hline\hline
Category & 1992 & 1994 & 1996/1998 & 2000 & 2002 & 2004/2005 & 2006 \\
\hline
Cereals              & A001--020 & A001--021 & A001--021 & A001--021 & A001--021 & A001--022 & A001--024 \\
Meat/fish/seafood    & A021--058 & A022--059 & A022--059 & A022--059 & A022--070 & A023--071 & A025--074 \\
Dairy + eggs         & A059--078 & A060--079 & A060--079 & A060--079 & A071--090 & A072--091 & A075--094 \\
Oils and fats        & A079--083 & A080--084 & A080--084 & A080--084 & A091--096 & A092--097 & A095--100 \\
Vegetables \& fruit  & A084--143 & A085--145 & A085--146 & A085--146 & A097--168 & A098--169 & A101--172 \\
Sugar and honey      & A144--146 & A146--148 & A147--149 & A147--149 & A169--171 & A170--172 & A173--175 \\
Soft drinks          & A185--189 & A188--193 & A189--194 & A189--194 & A212--218 & A212--218 & A212--218 \\
Desserts             & A174--181 & A177--183 & A178--184 & A178--184 & A201--206 & A202--206 & A202--206 \\
Coffee               & A147--153 & A149--155 & A150--156 & A150--156 & A172--178 & A173--179 & A173--179 \\
Spices               & A154--165 & A156--166 & A157--167 & A157--167 & A179--190 & A180--191 & A180--191 \\
Other food prep.     & A182      & A184--185 & A185--186 & A185--186 & A207--208 & A207--208 & A207--208 \\
Baby food            & A166--167 & A167--169 & A168--170 & A168--170 & A191--193 & A192--194 & A192--194 \\
Takeout (for home)   & A168--172 & A170--175 & A171--176 & A171--176 & A194--198 & A195--199 & A195--199 \\
Food eaten outside   & A199--202 & A204--207 & A205--208 & A206--209 & A235--239 & A235--239 & A243--247 \\
Tobacco              & A203--205 & A208--210 & A209--211 & A211--213 & A240--242 & A240--242 & A239--241 \\
Alcohol              & A190--198 & A194--203 & A195--204 & A195--204 & A219--234 & A219--234 & A219--234 \\
Pet food             & A183--184 & A186--187 & A187--188 & A187--188 & A210--211 & A210--211 & A210--211 \\
\hline\hline
\end{tabular}
\end{table}

\paragraph{Cross-wave catalogue corrections.}
Several code boundaries were corrected relative to prior versions of the code
based on verification against the official INEGI catalogue PDFs:

\begin{itemize}
  \item \textbf{Outside food, 1994}: starts at A204 (not A203), because A203
    in 1994 belongs to the tobacco section (\textit{Cigarros}). The 1992
    outside-food range (A199--A202) anchors the correct boundary; the 1994
    catalogue inserts one additional tobacco item, shifting outside food up by
    one.
  \item \textbf{Outside food and tobacco, 1996/1998 vs.\ 2000}: the catalogue
    shifts by one code between 1996/1998 and 2000, handled by a conditional
    \texttt{if inlist(`year', 1996, 1998)} block in the code (outside food:
    A205--A208 vs.\ A206--A209; tobacco: A209--A211 vs.\ A211--A213).
  \item \textbf{Vegetables and fruit, 2004/2005}: the upper bound is A169,
    which includes \textit{Jalea y mermelada} (jam/jelly), a processed fruit
    product. This item was correctly retained within \texttt{vegg\_fruit\_d}
    rather than being reclassified to the sugar category.
  \item \textbf{Packaged food, 2002+}: identified by two specific codes A209
    and A243 using \texttt{inlist(CLAVE, "A209", "A243")} rather than a range,
    because these two codes are non-contiguous in the catalogue.
\end{itemize}

%--------------------------------------------------------------------------
\subsection{Health Expenditure Variables}
\label{app:enigh_health}
%--------------------------------------------------------------------------

Health spending is extracted from the same \texttt{gastos.dta} file using
J-prefix catalogue codes. Six categories are constructed:

\begin{description}
  \item[\texttt{medical\_outpatient}] Outpatient consultations with doctors,
    dentists, nurses, and midwives; laboratory tests; X-rays; and related
    outpatient diagnostic services.
  \item[\texttt{drugs\_prescribed}] Prescription medications obtained through
    outpatient and inpatient channels.
  \item[\texttt{medical\_inpatient}] Hospital fees, surgical procedures,
    overnight stays, and inpatient diagnostic tests.
  \item[\texttt{drugs\_overcounter}] Over-the-counter medications and first-aid
    materials (bandages, antiseptics, analgesics without prescription).
  \item[\texttt{ortho}] Eyeglasses, contact lenses, dentures, hearing aids,
    prosthetics, orthopedic devices, and repairs of such devices.
  \item[\texttt{insurance\_cost}] Health insurance premiums and hospital
    membership fees paid out of pocket.
\end{description}

Table~\ref{tab:health_jcodes} reports the J-code ranges by wave. A key
correction involved \texttt{drugs\_overcounter}: the lower bound in
1992/1994 is J029 and in 1996/1998 is J033, rather than J028 and J032
respectively, because the preceding code in each case captures a different item
(surgical/dental appliances). The 2000 lower bound differs by one from
1996/1998 (J034 vs.\ J033), handled via a conditional block.

\begin{table}[h!]
\centering
\caption{Health Expenditure J-Code Ranges by ENIGH Wave}
\label{tab:health_jcodes}
\small
\begin{tabular}{llllll}
\hline\hline
Variable & 1992/1994 & 1996/1998 & 2000 & 2002 & 2004/2005/2006 \\
\hline
\texttt{medical\_outpatient} & J001--003, & J001--003, & same & J001--004 & J016--019, \\
                              & J005, J006, J009 & J005, J006, J009 & & & J036 \\
\texttt{drugs\_prescribed}   & J004, J011 & J004, J011 & same & J005--023 & J020--035, \\
                              &            &            &      &           & J037--038 \\
\texttt{medical\_inpatient}  & J010, J012--015 & same & same & J026--030 & J039--043 \\
\texttt{drugs\_overcounter}  & J029--036 & J033--038 & J034--038 & J048--065 & J044--059 \\
\texttt{ortho}               & J037--041 & J039--043 & same & J070--075 & J065--069 \\
\texttt{insurance\_cost}     & J042--043 & J044--045 & same & J076--077 & J070--072 \\
\hline\hline
\end{tabular}
\end{table}

%--------------------------------------------------------------------------
\subsection{Household Income Variables}
\label{app:enigh_income}
%--------------------------------------------------------------------------

Income data are extracted from \texttt{ingresos.dta}, which records individual
income sources identified by P-prefix codes. Each source is an indicator
variable (1 if the P-code belongs to that category, 0 otherwise), multiplied by
\texttt{ING\_TRI/3} to yield monthly amounts. Individual-level totals are
summed within person (\texttt{bys FOLIO NUM\_REN: egen total = total(aux)}) and
household-level totals are summed across all household members
(\texttt{bys FOLIO: egen total = total(aux)}).

The five income components are:

\begin{description}
  \item[\texttt{wage\_d}] Subordinate-employment income: wages, salaries,
    commissions, overtime, Christmas bonus (\textit{aguinaldo}), vacation
    premiums, and profit-sharing (\textit{reparto de utilidades}) paid by
    private employers, cooperatives, and enterprise partnerships.
  \item[\texttt{indep\_w\_d}] Self-employment income: net income from own
    agricultural, livestock, industrial, commercial, and service businesses.
    Does not include capital returns from cooperatives or enterprise
    memberships (those go to \texttt{capital\_d}).
  \item[\texttt{capital\_d}] Capital income: rental income from land,
    buildings, and machinery; interest; dividends; royalties; and net profits
    from cooperative and enterprise ownership (\textit{ganancias netas}).
  \item[\texttt{transfer\_d}] All transfers: pensions, severance pay,
    scholarships, government and NGO donations, family remittances, PROGRESA,
    and PROCAMPO. See Section~\ref{app:enigh_progresa} for the PROGRESA
    sub-component.
  \item[\texttt{other\_d}] Other current income: proceeds from the sale of
    second-hand goods and miscellaneous current receipts.
  \item[\texttt{financial\_d}] Financial/liquidity receipts: loans received,
    pension fund withdrawals, proceeds from sale of real estate or financial
    assets. \textit{Excluded from all income aggregates;} these capture
    balance-sheet events rather than current income flows.
\end{description}

Table~\ref{tab:income_pcodes} summarises the P-code mappings across waves.
Several boundaries required careful handling:

\begin{itemize}
  \item In 1992, P012 (\textit{venta de productos agr\'icolas}) and P014
    (\textit{venta de animales}) are farm sales revenue of uncertain
    classification and are excluded from all categories.
  \item In 2002, the code P023 was tentatively assigned to \texttt{indep\_w\_d}
    based on catalogue position; in 2004/2005/2006 the corresponding category
    was moved to \texttt{capital\_d} as cooperative/enterprise net profits were
    separated from own-business income.
  \item Cooperative and enterprise wages (codes P018 in 2002; P017, P019--P037
    in 2004--2006) are included in \texttt{wage\_d}, not \texttt{capital\_d},
    because they represent labor compensation rather than returns to ownership.
\end{itemize}

\begin{table}[h!]
\centering
\caption{Income P-Code Assignments by ENIGH Wave}
\label{tab:income_pcodes}
\small
\begin{tabular}{llllll}
\hline\hline
Variable & 1992 & 1994/1996 & 1998/2000 & 2002 & 2004/2005/2006 \\
\hline
\texttt{wage\_d}     & P001--006 & P001--005, P014 & P001--009, P018 & P001--009, & P001--009, \\
                     &           &                 &                 & P018, P020, P022 & P017, P019--037 \\
\texttt{indep\_w\_d} & P007--011, P013 & P006--013 & P010--017 & P010--017, P023 & P010--016 \\
\texttt{capital\_d}  & P015--021 & P015--022 & P019--027 & P019, P021, & P018, P028, P038, \\
                     &           &           &           & P024--036   & P039--047 \\
\texttt{transfer\_d} & P022--027 & P023--028, P043 & P028--034 & P037--047 & P048--060 \\
\texttt{other\_d}    & P028--029 & P029--030 & P035--036 & P048 & P061 \\
\texttt{financial\_d}& Q012, Q014$^*$ & P031--036 & P037--043 & P049--065 & P062--075 \\
\hline\hline
\multicolumn{6}{l}{\footnotesize $^*$In 1992, financial receipts were coded in the income file under Q-prefix codes rather than in a separate eroga file.}
\end{tabular}
\end{table}

\paragraph{Income aggregates.}
Two aggregate income measures are constructed after appending all waves:

\begin{align*}
\texttt{earnings} &= \texttt{wage} + \texttt{indep\_w} \\
\texttt{income}   &= \texttt{wage} + \texttt{indep\_w} + \texttt{capital}
                     + \texttt{transfer} + \texttt{other\_inc}
\end{align*}

available at both the individual (\texttt{\_ind}) and household (\texttt{\_hh})
level. The survey's own total-income aggregate (\texttt{INGMON} from
\texttt{concen.dta}) is retained as \texttt{hh\_income\_mxn\_tot} for
comparison; it is computed quarterly and divided by three, then rescaled for
1992. \texttt{financial\_d} is excluded from the constructed aggregate,
consistent with national accounts conventions.

%--------------------------------------------------------------------------
\subsection{Financial Outflows: Savings, Debt, and Currency}
\label{app:enigh_savings}
%--------------------------------------------------------------------------

Financial outflow variables are drawn from \texttt{eroga.dta} (available for
1994 and later; in 1992 the equivalent items appear under Q-prefix codes in
\texttt{ingresos.dta}) using Q-prefix catalogue codes. Monthly amounts are
\texttt{ERO\_TRI/3}. Four variables are constructed:

\begin{description}
  \item[\texttt{savings}] Deposits into savings accounts (Q001 in all waves).
  \item[\texttt{loans}] Loans made by the household to persons outside the
    household (Q002 in all waves).
  \item[\texttt{debt}] Credit-card payments, loan repayments, and mortgage
    payments. The codes differ slightly: Q003--Q004 plus Q010 for 1994--2002;
    Q003--Q005 plus Q011 for 2004--2006.
  \item[\texttt{currency}] Purchase of coins, precious metals, and foreign
    currency (Q005 for 1992--2002; Q006 for 2004--2006).
\end{description}

These variables capture household balance-sheet activity rather than current
income or expenditure. In the mechanisms regressions they test whether PROGRESA
affects household saving behaviour or debt repayment---channels that could link
the transfer to long-run welfare and mortality.

%--------------------------------------------------------------------------
\subsection{PROGRESA Transfer Variables}
\label{app:enigh_progresa}
%--------------------------------------------------------------------------

This section documents the construction of two distinct variables used to
capture PROGRESA exposure in the ENIGH: (i) the \textit{income amount} of
PROGRESA transfers received by a household or individual (\texttt{progresa\_hh},
\texttt{progresa\_ind}), and (ii) a binary \textit{beneficiary flag} indicating
whether any household member holds a PROGRESA scholarship
(\texttt{progresa\_benef\_hh}, \texttt{progresa\_benef\_ind}). Both variables
are used as \textit{outcome variables} in the mechanisms analysis; they are
\textit{not} used to assign program treatment, which is drawn from
administrative enrollment records (see \texttt{aamr\_011326.do} and
Section~\ref{sec:data}). The purpose of including them as outcomes is to verify
that the program's cash transfers are credibly visible in the household survey
for the treated municipalities, providing a survey-based check on program reach.

\subsubsection{Survey Instrument Evolution}
\label{app:enigh_evolution}

Table~\ref{tab:enigh_pcodes} summarises how PROGRESA/Oportunidades transfers
are recorded in the ENIGH income questionnaire across waves.

\begin{table}[h!]
\centering
\caption{ENIGH P-Codes for PROGRESA/Oportunidades Transfers, by Wave}
\label{tab:enigh_pcodes}
\small
\begin{tabular}{lllll}
\hline\hline
Wave & PROGRESA code & Govt.\ transfers & Non-govt.\ transfers & Notes \\
\hline
1992--1996 & --- & P025/P026 & P026/P027 & Program did not exist \\
1998       & P031 (implicit) & P031 & P032 & Bundled with all institutional grants \\
2000       & P031 (explicit) & P031 & P032 & SEDESOL annotation in enumerator manual \\
2002       & P046 (own code) & P043 & P042 & First dedicated PROGRESA code \\
2004/2005  & P059 (own code) & P056 & P055 & Fully disaggregated \\
2006       & P059 (own code) & P056 & P055 & Same as 2004/2005 \\
\hline\hline
\end{tabular}
\end{table}

\noindent
The 2000 ENIGH enumerator manual (\textit{Manual del Entrevistador}, INEGI
2000) contains an explicit SEDESOL annotation for P031:

\begin{quote}
\textit{``La Secretar\'ia de Desarrollo Social (SEDESOL) ha implementado un
mecanismo de ayuda para las familias de escasos recursos (Programa Progresa).
Cuando los ingresos monetarios sean otorgados por este programa el cual es un
apoyo econ\'omico que se proporciona en efectivo en la modalidad de beca o
donativo, se registran en este rengl\'on al miembro del hogar responsable del
menor.''}
\end{quote}

This confirms that PROGRESA transfers were directed to P031 for the 2000 wave.
The same P-code structure applied in 1998, though the 1998 manual did not yet
contain this annotation.

\subsubsection{Income Amount: Waves 1992--1996}

PROGRESA did not exist prior to 1997. \texttt{progresa\_d = 0} for all
pre-program waves; \texttt{progresa\_ind} and \texttt{progresa\_hh} are zero
for all observations in 1992, 1994, and 1996.

\subsubsection{Income Amount: Waves 2002--2006}

Starting in 2002, PROGRESA/Oportunidades has a dedicated P-code (P046 in 2002;
P059 in 2004--2006). The income variable is constructed directly from these
codes with no imputation. This is the cleanest identification of PROGRESA
amounts in the ENIGH.

\subsubsection{Income Amount: Waves 1998 and 2000 --- Imputation}
\label{app:enigh_imputation}

For 1998 and 2000, P031 aggregates PROGRESA transfers together with all other
government institutional scholarships and grants. There is no way to
disentangle them at the respondent level. We use the following imputation,
applied after appending all waves:

\begin{equation}
\widehat{\texttt{progresa}}_{i,t} =
  \bigl(\texttt{benef\_don\_gob}_{i,t}
    + \texttt{benef\_don\_non\_gob}_{i,t}\bigr)
  \times \hat{\omega}_{2002}, \quad t \in \{1998, 2000\}
\label{eq:progresa_share}
\end{equation}

where $\hat{\omega}_{2002}$ is PROGRESA's share of the combined institutional
transfer pool in 2002, computed from national sample means:

\begin{equation}
\hat{\omega}_{2002} =
  \frac{\overline{\texttt{progresa}}_{2002}}
       {\overline{\texttt{progresa}}_{2002}
        + \overline{\texttt{benef\_don\_gob}}_{2002}
        + \overline{\texttt{benef\_don\_non\_gob}}_{2002}}
\label{eq:share_2002}
\end{equation}

The Stata implementation:

\begin{verbatim}
sum benef_don_non_gob_ind if year == 2002
local m_benef_non_gob = `r(mean)'
sum benef_don_gob_ind   if year == 2002
local m_benef_gob = `r(mean)'
sum progresa_ind        if year == 2002
local m_progresa = `r(mean)'

local share_2002 = `m_progresa' / ///
    (`m_progresa' + `m_benef_gob' + `m_benef_non_gob')

replace progresa_ind = ///
    (benef_don_non_gob_ind + benef_don_gob_ind) * `share_2002' ///
    if inlist(year, 1998, 2000)
\end{verbatim}

\paragraph{Alternative proxies.}
Two robustness alternatives are also constructed and retained in the dataset:

\begin{description}
  \item[Alt1: State-level share, same pool.] $\hat{\omega}_{e,2002}$ is
    computed separately for each state $e$ using state-specific means; falls
    back to the national mean for states with no positive 2002 observations.
    Captures geographic variation in PROGRESA rollout without the noise of
    municipality-level small samples.
  \item[Alt2: State-level share, government-transfers pool only.] Restricts the
    denominator to P043 (government institutional transfers) in 2002 and
    scales only \texttt{benef\_don\_gob\_ind} (P031) in 1998/2000, excluding
    the non-governmental channel P032. Theoretically cleaner because the SEDESOL
    annotation directs PROGRESA to the government channel only.
\end{description}

\paragraph{Identifying assumptions and limitations.}
The imputation rests on three assumptions, each with caveats:

\begin{enumerate}
  \item \textit{Stable composition.} PROGRESA's fraction of institutional
    transfers in 2002 proxies its fraction in 1998/2000. Plausible in highly
    marginalized rural municipalities, where PROGRESA dominated the
    institutional transfer category.
  \item \textit{No over-reporting bias in 2002.} \citet{hernandez2007}
    document that Oportunidades income in the 2002 ENIGH grew by more than 100
    percent relative to 2000, versus only 59 percent in administrative budget
    data, attributable in part to SEDESOL oversampling of beneficiary
    households. If 2002 PROGRESA income is over-reported, $\hat\omega_{2002}$
    is inflated and imputed 1998/2000 amounts are biased upward in levels.
    Within-municipality fixed effects in our DiD estimator absorb any
    level bias; the concern is relevant only for cross-wave levels comparisons.
  \item \textit{No contamination from non-PROGRESA transfers.} P031 also
    includes CONAFE stipends and other institutional grants. In our sample of
    highly marginalized municipalities, PROGRESA was the dominant government
    transfer programme, making contamination likely small.
\end{enumerate}

\subsubsection{Beneficiary Flag}
\label{app:enigh_benef_flag}

The binary beneficiary indicator (\texttt{progresa\_benef\_ind}) comes from
the population file \texttt{pob.dta} and captures whether an individual holds a
current PROGRESA scholarship. Its construction differs by wave:

\begin{table}[h!]
\centering
\caption{Construction of \texttt{progresa\_benef\_ind} by ENIGH Wave}
\label{tab:benef_flag}
\small
\begin{tabular}{lll}
\hline\hline
Wave & Population variable & Condition \\
\hline
1992--1996 & --- & = 0 (program did not exist) \\
1998       & --- & = 0 (no PROGRESA entry in \texttt{pob.dta}) \\
2000       & \texttt{PROP\_BECA} & = 1 if \texttt{PROP\_BECA == 5}
             (\textit{Becas de PROGRESA}) \\
2002       & \texttt{BECA}       & = 1 if \texttt{BECA == 5} \\
2004       & \texttt{BECAS}      & = 1 if \texttt{BECAS == 1} \\
2005/2006  & \texttt{BECA}       & = 1 if \texttt{BECA == 1} \\
\hline\hline
\end{tabular}
\end{table}

\noindent
The household flag equals one if any member satisfies the individual condition:
\texttt{bys FOLIO: egen progresa\_benef\_hh = max(progresa\_benef\_ind)}.

Two coverage limitations apply. First, the 1998 population file does not
contain a scholarship-type variable identifying PROGRESA recipients despite the
program being active; \texttt{progresa\_benef\_ind = 0} for all 1998
observations. Second, \texttt{PROP\_BECA == 5} in 2000 identifies only the
\textit{educational} scholarship component---households with a school-age child
receiving beca PROGRESA. Nutrition and health transfers disbursed to mothers
without school-age children are not captured, making the 2000 flag a lower
bound on true beneficiary status. Since both variables are used as outcomes
rather than treatment indicators, these limitations affect our ability to
\textit{detect} program reach in the survey, not the validity of the
causal identification.

\subsubsection{Role in the Empirical Analysis}

\texttt{progresa\_hh} and \texttt{progresa\_benef\_hh} appear in the mechanisms
table as dependent variables to verify that program cash is visible in the
survey among households in high-intensity municipalities. Measurement error
in these outcome variables inflates standard errors but does not bias the
difference-in-differences coefficients, which are identified from
within-municipality variation absorbed by municipality fixed effects.

%--------------------------------------------------------------------------
\subsection{Deflation and Real Value Construction}
\label{app:enigh_deflation}
%--------------------------------------------------------------------------

After all waves are appended, every nominal monetary variable is converted to
real 2025 USD using a wave-specific conversion factor:

\begin{equation}
\text{real}_{i,t} = \text{nominal}_{i,t} \times
  \frac{\text{CPI}_{2025}}{\text{CPI}_{t}} \times
  \frac{1}{\text{MXN/USD}_{2025}}
\end{equation}

The Mexican CPI series (base July 2018 = 100) is evaluated at September of
each survey year, matching the typical ENIGH fieldwork period. The exchange
rate is the September 2025 MXN/USD rate. The conversion factors used are:

\begin{table}[h!]
\centering
\caption{CPI and Exchange-Rate Deflators by Wave}
\label{tab:deflators}
\small
\begin{tabular}{lrr}
\hline\hline
Wave & CPI (Sep, base July 2018 = 100) & Combined factor (real 2025 USD / MXN nominal) \\
\hline
1992 & 12.443 & $\times (141.197/12.443) \times (1/20.52)$ \\
1994 & 14.536 & $\times (141.197/14.536) \times (1/20.52)$ \\
1996 & 27.113 & $\times (141.197/27.113) \times (1/20.52)$ \\
1998 & 37.327 & $\times (141.197/37.327) \times (1/20.52)$ \\
2000 & 47.061 & $\times (141.197/47.061) \times (1/20.52)$ \\
2002 & 52.422 & $\times (141.197/52.422) \times (1/20.52)$ \\
2004 & 57.298 & $\times (141.197/57.298) \times (1/20.52)$ \\
2005 & 59.309 & $\times (141.197/59.309) \times (1/20.52)$ \\
2006 & 61.737 & $\times (141.197/61.737) \times (1/20.52)$ \\
\hline\hline
\end{tabular}
\end{table}

\noindent
The 1992 wave additionally requires a pre-deflation rescaling by 1/1,000 to
convert old Mexican pesos to new pesos (redenomination of January 1993).
For food variables, per-capita versions are generated as \texttt{food / n\_hh}
and natural-log versions as \texttt{ln(food\_pc)}. For income variables,
natural logs are generated as \texttt{ln(income)}. All MXN-denominated
originals are retained with the suffix \texttt{\_mxn}.

%--------------------------------------------------------------------------
\subsection{Derived Behavioral Indices}
\label{app:enigh_derived}
%--------------------------------------------------------------------------

After deflation, three families of derived variables are constructed from the
real-valued food, health, and financial outflow variables. All monetary inputs
are in real 2025 USD. Ratios are set to missing when the denominator is zero
to avoid both arithmetic artifacts and economically meaningless observations
(e.g., an OTC-to-Rx ratio for a household that never purchased prescription
drugs). These variables are constructed after the income and food deflation
loops in \texttt{01\_enigh\_data.do}, immediately before the final
\texttt{save}.

\subsubsection{Diet Quality Indices}
\label{app:enigh_diet}

PROGRESA's nutrition component aimed to shift consumption away from
calorie-dense staple grains toward protein and micronutrient-rich foods
\citep{hoddinott2008,leroy2008}. Four share variables and one count index
capture this potential dietary shift:

\begin{align}
\texttt{protein\_share}    &= \frac{\texttt{meat\_dairy}}{\texttt{food\_exp}}
    \label{eq:protein_share} \\[4pt]
\texttt{staples\_share}    &= \frac{\texttt{cereals}}{\texttt{food\_exp}}
    \label{eq:staples_share} \\[4pt]
\texttt{vegg\_fruit\_share}&= \frac{\texttt{vegg\_fruit}}{\texttt{food\_exp}}
    \label{eq:vegg_share} \\[4pt]
\texttt{unhealthy\_share}  &= \frac{\texttt{sugar\_fat\_drink}
    + \texttt{alcohol} + \texttt{tobacco}}{\texttt{food\_exp}}
    \label{eq:unhealthy_share}
\end{align}

\noindent
All four are defined only for households with $\texttt{food\_exp} > 0$.
\texttt{food\_exp} is the sum of cereals, meat/dairy, vegg/fruit,
sugar/fat/drink, coffee/spices/other, outside food, and packaged and baby food
(see Section~\ref{app:enigh_food}).

The fifth index is a dietary diversity count:
\begin{equation}
\texttt{diversity\_index} = \sum_{c \in \mathcal{C}} \mathbf{1}[\texttt{spend}_c > 0]
\label{eq:diversity}
\end{equation}
where $\mathcal{C} = \{\text{cereals, meat/dairy, vegg/fruit, sugar/fat/drink,
coffee/spices/other, outside food, alcohol}\}$, giving a score of 0–7.
Set to missing if \texttt{food\_exp} = 0.

\paragraph{Interpretation.}
\texttt{staples\_share} is a food insecurity proxy: households under chronic
caloric scarcity anchor their diet on cheap, filling grains. A PROGRESA effect
would manifest as a \emph{fall} in \texttt{staples\_share} and a
\emph{rise} in \texttt{protein\_share}, \texttt{vegg\_fruit\_share}, and
\texttt{diversity\_index} — directly mapping onto reduced micronutrient
deficiency and protein-energy malnutrition, pathways linked to lower
infectious disease mortality and reduced frailty in the elderly
\citep{leroy2008,barham2013}.

\subsubsection{Healthcare Substitution Ratios}
\label{app:enigh_healthcare_ratios}

PROGRESA conditionality required bimonthly health clinic visits; the
behavioral prediction is a shift from self-medication toward formal care.
Three variables test different facets of this channel:

\begin{align}
\texttt{rx\_to\_visit\_ratio}  &= \frac{\texttt{drugs\_prescribed}}
    {\texttt{medical\_outpatient}},
    \quad \text{if } \texttt{medical\_outpatient} > 0
    \label{eq:rx_visit} \\[4pt]
\texttt{otc\_to\_rx\_ratio}    &= \frac{\texttt{drugs\_overcounter}}
    {\texttt{drugs\_prescribed}},
    \quad \text{if } \texttt{drugs\_prescribed} > 0
    \label{eq:otc_rx} \\[4pt]
\texttt{health\_share}         &= \frac{\texttt{health\_exp}}
    {\texttt{hh\_expenditure}},
    \quad \text{if } \texttt{hh\_expenditure} > 0
    \label{eq:health_share}
\end{align}

\noindent
\texttt{health\_exp} $= \texttt{medical\_outpatient} + \texttt{drugs\_prescribed}
+ \texttt{medical\_inpatient} + \texttt{drugs\_overcounter} + \texttt{ortho}$
(see Section~\ref{app:enigh_health}).

\paragraph{Interpretation.}
$\texttt{rx\_to\_visit\_ratio}$ captures whether outpatient visits translate
into prescriptions. A \emph{rise} indicates that formal medical encounters
produce more follow-through care — consistent with PROGRESA clinic visits
increasing diagnostic depth. $\texttt{otc\_to\_rx\_ratio}$ captures the
self-medication intensity: if conditionality shifts households from pharmacy
runs to clinic visits, this ratio should \emph{fall} post-program. A pure
income effect without conditionality would not necessarily change
$\texttt{otc\_to\_rx\_ratio}$, making this ratio a useful diagnostic to
separate the conditionality channel from the income channel.
$\texttt{health\_share}$ tests whether health receives a larger fraction of
the household budget under PROGRESA — an overall budget-allocation channel.

\subsubsection{Financial Stress Composite}
\label{app:enigh_fin_stress}

Income uncertainty reduction is a plausible indirect mortality channel
through allostatic load \citep{aizer2016}: households with lower debt burdens
and positive saving positions face less chronic financial stress, with downstream
effects on cardiovascular and stress-related mortality. Two variables proxy
financial stress:

\begin{align}
\texttt{net\_fin\_position} &= \texttt{savings} - \texttt{debt}
    \label{eq:net_fin} \\[4pt]
\texttt{debt\_to\_income}   &= \frac{\texttt{debt}}{\texttt{hh\_earnings}},
    \quad \text{if } \texttt{hh\_earnings} > 0
    \label{eq:debt_income}
\end{align}

\noindent
\texttt{savings} and \texttt{debt} are monthly real 2025 USD amounts from
\texttt{eroga.dta} (see Section~\ref{app:enigh_savings}).
\texttt{hh\_earnings} = \texttt{wage\_hh} + \texttt{indep\_w\_hh} captures
own-labor income only, excluding transfers, so the denominator is independent
of the PROGRESA transfer itself — the ratio measures the burden of debt service
relative to the household's own productive capacity.

\paragraph{Interpretation.}
A positive treatment effect on \texttt{net\_fin\_position} (savings
accumulation net of debt service) and a negative effect on
\texttt{debt\_to\_income} (debt burden relative to earnings) would support
the financial stress relief channel. These patterns are informative regardless
of whether income itself rises, because PROGRESA transfers could allow
households to retire debt even without changing earnings.

\begin{table}[h!]
\centering
\caption{Summary of Derived Behavioral Indices}
\label{tab:derived_indices}
\small
\begin{tabular}{llll}
\hline\hline
Variable & Formula & Defined when & Predicted direction \\
\hline
\texttt{protein\_share}    & meat\_dairy / food\_exp     & food\_exp $> 0$ & $\uparrow$ \\
\texttt{staples\_share}    & cereals / food\_exp          & food\_exp $> 0$ & $\downarrow$ \\
\texttt{vegg\_fruit\_share}& vegg\_fruit / food\_exp      & food\_exp $> 0$ & $\uparrow$ \\
\texttt{unhealthy\_share}  & (sugar+fat+alc+tob) / food\_exp & food\_exp $> 0$ & ambiguous \\
\texttt{diversity\_index}  & $\sum \mathbf{1}[\text{spend}_c > 0]$ & food\_exp $> 0$ & $\uparrow$ \\[3pt]
\texttt{rx\_to\_visit\_ratio} & drugs\_Rx / outpatient  & outpatient $> 0$ & $\uparrow$ \\
\texttt{otc\_to\_rx\_ratio}   & drugs\_OTC / drugs\_Rx  & Rx $> 0$         & $\downarrow$ \\
\texttt{health\_share}        & health\_exp / hh\_exp   & hh\_exp $> 0$    & $\uparrow$ \\[3pt]
\texttt{net\_fin\_position} & savings $-$ debt          & always           & $\uparrow$ \\
\texttt{debt\_to\_income}   & debt / hh\_earnings       & earnings $> 0$   & $\downarrow$ \\
\hline\hline
\multicolumn{4}{l}{\footnotesize Predicted direction: under PROGRESA income + conditionality effects.}\\
\multicolumn{4}{l}{\footnotesize ``ambiguous'': income effect could raise spending on sugar/fat; conditionality unlikely to change it directly.}
\end{tabular}
\end{table}

%--------------------------------------------------------------------------
\subsection{Literature Context}
\label{app:enigh_lit}
%--------------------------------------------------------------------------

The challenge of identifying PROGRESA in pre-2002 ENIGH waves is documented
by \citet{hernandez2007}, who show that the number of income questions grew from
36 in 2000 to 48 in 2002 to 61 in 2004, and that SEDESOL partially funded the
2002 ENIGH to achieve representativeness among Oportunidades beneficiaries.
Researchers requiring clean PROGRESA identification typically adopt one of three
strategies: (i) use the PROGRESA experimental evaluation surveys (ENCEL/ENCASEH)
\citep{gertler2004,parker2005,skoufias2001}; (ii) merge administrative enrollment
records at the municipality level and bypass the survey altogether
\citep{barham2013}; or (iii) restrict ENIGH-based analysis to 2002 and later
\citep{parker2008}. Our approach---using 1998/2000 with the share imputation as
a \textit{dependent} variable check on survey-based program reach---is most
similar in spirit to strategy (ii), with the ENIGH serving as a validity check
on the administrative treatment variable rather than as an identification source.
\citet{barham2013} is the closest existing study to our design: a
municipality-level DiD exploiting PROGRESA administrative enrollment records and
vital statistics to estimate effects on elderly mortality.

%--------------------------------------------------------------------------
% REFERENCES (add to main .bib file)
%--------------------------------------------------------------------------
% @article{hernandez2007,
%   author  = {Hern\'andez Meza, Gonzalo and others},
%   title   = {Los problemas de comparabilidad de las {ENIGH} y su efecto
%              en la medici\'on de la pobreza},
%   journal = {Papeles de Poblaci\'on},
%   year    = {2007},
%   number  = {51},
%   pages   = {95--125}
% }
% @article{gertler2004,
%   author  = {Gertler, Paul},
%   title   = {Do Conditional Cash Transfers Improve Child Health?},
%   journal = {American Economic Review},
%   year    = {2004},
%   volume  = {94},
%   number  = {2},
%   pages   = {336--341}
% }
% @article{parker2005,
%   author  = {Parker, Susan W. and Teruel, Graciela M.},
%   title   = {Randomization and Social Program Evaluation: The Case of {Progresa}},
%   journal = {The Annals of the American Academy of Political and Social Science},
%   year    = {2005},
%   volume  = {599},
%   pages   = {199--219}
% }
% @article{skoufias2001,
%   author  = {Skoufias, Emmanuel and Parker, Susan W. and Behrman, Jere R.
%              and Pessino, Carola},
%   title   = {Conditional Cash Transfers and Their Impact on Child Work and Schooling},
%   journal = {Econom\'ia},
%   year    = {2001},
%   volume  = {2},
%   number  = {1},
%   pages   = {45--96}
% }
% @article{barham2013,
%   author  = {Barham, Tania and Rowberry, Jacob},
%   title   = {Living longer: The effect of the {Mexican} conditional cash
%              transfer program on elderly mortality},
%   journal = {Journal of Development Economics},
%   year    = {2013},
%   volume  = {105},
%   pages   = {226--236},
%   doi     = {10.1016/j.jdeveco.2013.08.002}
% }
% @article{hoddinott2008,
%   author  = {Hoddinott, John and Bassett, Lucy},
%   title   = {Conditional Cash Transfer Programs and Nutrition in {Latin America}:
%              Assessment of Impacts and Strategies for Improvement},
%   journal = {IFPRI Discussion Paper},
%   year    = {2008},
%   number  = {00708}
% }
% @article{leroy2008,
%   author  = {Leroy, Jef L. and Ruel, Marie and Verhofstadt, Ellen},
%   title   = {The impact of conditional cash transfer programmes on child
%              nutrition: a review of evidence using a programme theory framework},
%   journal = {Journal of Development Effectiveness},
%   year    = {2009},
%   volume  = {1},
%   number  = {2},
%   pages   = {103--129}
% }
% @article{aizer2016,
%   author  = {Aizer, Anna and Eli, Shari and Ferrie, Joseph and Lleras-Muney, Adriana},
%   title   = {The Long-Run Impact of Cash Transfers to Poor Families},
%   journal = {American Economic Review},
%   year    = {2016},
%   volume  = {106},
%   number  = {4},
%   pages   = {935--971}
% }
% @incollection{parker2008,
%   author    = {Parker, Susan W. and Rubalcava, Luis and Teruel, Graciela M.},
%   title     = {Evaluating conditional schooling-health transfer programs},
%   booktitle = {Handbook of Development Economics},
%   year      = {2008},
%   volume    = {4},
%   pages     = {3963--3035}
% }
%===========================================================================
